---
title: "Homework 5"
author: "Group I: Demirbilek, Taroni, Zagatti"
date: "Spring 2020"
output:
  html_document:
    toc: no
header-includes:
- \usepackage{color}
- \definecolor{Purple}{HTML}{911146}
- \definecolor{Orange}{HTML}{CF4A30}
- \setbeamercolor{alerted text}{fg=Orange}
- \setbeamercolor{frametitle}{bg=Purple}
institute: University of Udine & University of Trieste
graphics: yes
fontsize: 10pt
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center', warning=FALSE, message=FALSE, fig.asp=0.625, dev='png', global.par = TRUE, dev.args=list(pointsize=10), fig.path = 'figs/', fig.height = 10, fig.width = 10)
```

```{r setup, include=FALSE}
library(MASS)
library(knitr)
local({
  hook_plot = knit_hooks$get('plot')
  knit_hooks$set(plot = function(x, options) {
    paste0('\n\n----\n\n', hook_plot(x, options))
  })
})
```

# {.tabset}

## DAAG {.tabset}


### Exercise 4.21


### Exercise 7.2

**Use anova() to compare the two models: 
`roller.lm <- lm(depression ~ weight, data = roller)`, `roller.lm2 <- lm(depression ~ weight + I(weigth^2), data = roller)`
Is there any justification for including the squared term?**


*solution :* 

```{r basic 1, echo=TRUE,warning=FALSE}
library(DAAG)
#help(roller)
roller.lm <- lm(depression ~ weight, data = roller)
roller.lm2 <- lm(depression ~ weight + I(weight^2), data = roller)
anova(roller.lm , roller.lm2)

```
These two models differ only in the use of `weight^2`. So ANOVA will test whether or not including this leads significant improvement. Adding this term spawned non-significant result (p=0.259). Thus we should reject second model, in other words adding `weight^2`didn't lead better result.

## CS {.tabset}


### Exercise 3.2



### Exercise 4.4

**Suppose that you have $n$ independent measurements of times between major aircraft disasters, $t_i$ , and believe that the probability density function for the $t_i$’s is of the form: $f(t) = ke^{−\lambda t^2}$ $t \geq 0$ where $λ$ and $k$ are the same for all $i$.**

**(a) By considering normal p.d.f., show that  $k = \sqrt{4 \lambda/\pi}.$**


*solution (a):*

$t \geq 0$ and $t = (x - \mu)$ so $x \geq \mu$. Since we consider only positive (right side) of normal p.d.f, we need to multiply it with by 2 to match the p.d.f $ke^{-\lambda t^2}$. So after this process we have:

$$f(t) = \frac{2}{\sigma\sqrt{2\pi}}\exp\{-\frac{(x - \mu)^2}{2\sigma^2}\} = ke^{-\lambda t^2}$$


$k = \frac{2}{\sigma\sqrt{2\pi}}$ , $\lambda = \frac{1}{2\sigma^2}$. Here we can find that

$$
k = \frac{2}{\sqrt{2\pi \sigma^2}} = \sqrt{\frac{4}{2\pi \sigma^2}} = \sqrt{\frac{4\lambda}{\pi}}
$$

**(b) Obtain a maximum likelihood estimator for $\lambda$.**

*solution (b):*

Likelihood of the given expression is the following

$$
L(t) = \prod_{i=1}^{n}ke^{−\lambda t_i^2}
$$
The log likelihood function carries the same information of the likelihood function, but it is mor manageable so for simplicity we are going to use log likelihood.

$$
\ell(t) = n \log(k) -\lambda \sum_{i=1}^{n}t_i^2 = n \log(\sqrt{4\lambda/ \pi} ) -\lambda \sum_{i=1}^{n}t_i^2 
$$
To find maximum likelihood estimator of $\lambda$ we need to derivate the log likelihood wit respect to $\lambda$ and find the $\lambda$ by setting equation zero.

$$
\frac{d\ell}{d\lambda} = \frac{n}{2\lambda}-\sum_{i=1}^{n}t_i^2 \\
\lambda = \frac{n}{2\sum_{i=1}^{n}t_i^2}
$$


**(c) Given observation of $T_i$ (in days) of 243, 14, 121, 63, 45, 407 and 34  use a generalised likelihood ratio test to
test $H_0 : \lambda = 10^{-4}$ against the alternative of no restriction on $\lambda$ at the $5%$ significance level. Note that if $V \sim X_1^2$ then $Pr[V \leq 3.841] = 0.95$.**

*solution (c):*

Likelihood test statistic is given by $W(\lambda_0) = 2(\ell(\hat{\lambda})-\ell(\lambda_0))$

```{r echo=TRUE,warning=FALSE}
log_lik_aircraft <- function(data,lambda){
  length(data) * log(sqrt(4*lambda/pi)) - lambda * sum(data^2)
}

data <-  c(243, 14, 121, 63, 45, 407 ,34)
mle <- length(data) / (2 * sum(data^2))
cat("Maximum likelihood estimation for lambda:",mle)

lambda_0 <- log_lik_aircraft(data,10^-4)
# since there is no restriction on lambda mle is chosen
lambda_mle <- log_lik_aircraft(data,mle)

lrt <- 2 * (lambda_mle - lambda_0)
# to find p-value by using Pr(W >= w_obs)
cat("p-value for test is:",pchisq(lrt,df=1,lower.tail = FALSE))

```
We have very small p-value, which indicates there is strong evidence against the null hypothesis.

## Bayesian {.tabset}

### BC 2.5
